---
title: "research"
output: html_document
date: "2024-05-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Preambula
Numerous studies has been dedicated to the area of brand-naming. Some of the researches are trying to discover the correlation between brand names likelihood and the sounds they use. Such attempt to map all concepts to specific sounds, is called sound mapping.
In the following research:

Trninić, Z. (2023). Creating effective brand names with sound symbolic mappings. Strategic Management-International Journal of Strategic Management and Decision Support Systems in Strategic Management, 28(2).

author has been investigated that different phonemes are associated with different qualities of objects. For example, active articulators such as /p/, /b/, /m/ are uttered more frequently when it comes to large objects.
Despite the fact that results of the paper's investigation indeed justify the hypothesis, the methodology is quit debatable. The most fallible part if presumption is that chosen products i.e. razors and soaps are associated with nothing more but sharpness ans softness. The problem is that razor are actually might be associated with softness, since for some people it is desirable to not get a scratch from using it.
The aim of this paper is to conduct similar experiment but with notion that the way people associate products is relative to each respondent.
Hence, the design of new experiment is the following:
1. For each respondent we generate 6 unique combinations of words. The very words' structure is CV-CV-CV, where each CV is from common categories.
2. Then we ask to give an association for chosen word
3. and at last we ask to chose the word again

# Data Description
Main data:
user_id - simply an id of respondent
firstly_chosen_group - the category of a word that was chosen by a user from the start of survey
associations - chosen word's associations that user came up with
image - a descriptive image of chosen word that user came up with
is_interested - yes/no answer to the question: would you be interested in a product simply because of the name you chose
lastly_chosen_group - the category of a word that was chosen by a user at the end of survey

features data:
transformed main data where associations and image column's values were used to calculate their own phonetic group stats


# libraries

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```


# Uploading data


## main data

```{r}
data <- read.csv(file="https://raw.githubusercontent.com/XvKuoMing/PhonoSymbolism/main/responses.csv", 
                 head=TRUE,
                 sep=",",
                 stringsAsFactors = TRUE
                 )
data
```


## support data

```{r}
features <- read.csv("https://raw.githubusercontent.com/XvKuoMing/PhonoSymbolism/main/responses_with_features.csv",
                     head=TRUE,
                     sep=',',
                     stringsAsFactors = TRUE)
features$lastly_chosen_group <- as.integer(features$lastly_chosen_group)
features$firstly_chosen_group <- as.integer(features$firstly_chosen_group)
features$is_interested <- as.integer(features$is_interested)
features
```


# Basic stats

## main data stats

```{r}
# create long pivot format
order_group_data <-
data %>%
  select(firstly_chosen_group, lastly_chosen_group) %>%
  pivot_longer(
    cols=c("firstly_chosen_group", "lastly_chosen_group"),
    names_to="order",
    values_to="group"
  )
order_group_data
```


```{r}
# group distribution for each order
ggplot(data=order_group_data,
       aes(x=group, group=order, fill=order)) +
  geom_density(alpha = 0.5) +
  scale_x_discrete(guide = guide_axis(angle = 90))
```


```{r}
# other way to print distribution
ggplot(data=order_group_data,
       aes(x=group,
           color=order,
           fill=order
           )
       ) +
  geom_bar(position="dodge") +
  scale_x_discrete(guide = guide_axis(angle = 90))

```


```
It seems that first and last groups are distributed equally. However, we need to check their contingency
```


```{r}
# contingency table
data.tabs <-
data %>%
  select(firstly_chosen_group, lastly_chosen_group) %>%
  table()
data.tabs
```

```{r}
# heatmap visualization
prop.table(data.tabs) %>%
  heatmap(margins=c(14, 7))
```

```
active_artucalor+back & upper_teeth:back not only very popular choices, but also repetitive
```


```{r}
# let's calculate the accuracy
equal_responses <- sum(as.character(data$firstly_chosen_group) == as.character(data$lastly_chosen_group))
accuracy <- equal_responses / nrow(data)

cat(accuracy * 100, "%")
```


```{r}
# before now we calculated the whole phonetically combination. However, vowels and consonants individually might have influence on choices 
data.separated <-
data %>%
  select(user_id, is_interested, firstly_chosen_group, lastly_chosen_group) %>%
  separate(firstly_chosen_group, c("firstly_con", "firstly_vow"), sep='\\+') %>%
  separate(lastly_chosen_group, c("lastly_con", "lastly_vow"), sep='\\+')

equal_responses_con <- sum(as.character(data.separated$firstly_con) == as.character(data.separated$lastly_con))
accuracy_con <- equal_responses_con / nrow(data.separated)
equal_responses_vow <- sum(as.character(data.separated$firstly_vow) == as.character(data.separated$lastly_vow))
accuracy_vow <- equal_responses_vow / nrow(data.separated)

cat("Accuracy of consonants repetition", accuracy_con, "\n")
cat("Accuracy of vowels repetition", accuracy_vow, "\n")
cat("Mean accuracy", (accuracy_con + accuracy_vow) / 2, "\n")
```


```{r}
# let's see the most popular consonants
data.separated.cons_tab <-
data.separated %>%
  select(firstly_con, lastly_con) %>%
  table()
data.separated.cons_tab
```


```{r}
# do the same for vowels
data.separated.vow_tab <-
data.separated %>%
  select(firstly_vow, lastly_vow) %>%
  table()
data.separated.vow_tab
```


## support data stats



```{r}
# beside our phonetics group there might be some influence from other variables, such as interest.
# For example, user may chose a phonetic group that he himself is not interested to. It would break the main idea of experiment i.e. repetition on group choices indicates the preferability of group as commercial product. Hence, we need to check data for such cases
data.interest <-
  data %>%
  select(firstly_chosen_group, is_interested) %>%
  table()
data.interest
```

```{r}
ggplot(data=data,
       aes(x=firstly_chosen_group,
           color=is_interested,
           fill=is_interested
           )
       ) +
  geom_bar(position="dodge") +
  scale_x_discrete(guide = guide_axis(angle = 90))
```


```{r}
# do the same for vowels and consonants, starting with vowels
data.separated %>%
  select(firstly_vow, is_interested) %>%
  ggplot(aes(x=firstly_vow,
             color=is_interested,
             fill=is_interested
           )
         ) +
  geom_bar(position="dodge") +
  scale_x_discrete(guide = guide_axis(angle = 90))
```


```{r}
data.separated %>%
  select(firstly_con, is_interested) %>%
  ggplot(aes(x=firstly_con,
             color=is_interested,
             fill=is_interested
           )
         ) +
  geom_bar(position="dodge") +
  scale_x_discrete(guide = guide_axis(angle = 90)) # we can see that positive answer is dominant
```



```{r}
# Now, let's see if there is any influence of chosen word to association
# firstly, print summary stats for count of each phonetic group in association+image
features %>%
  select(soft_palate, back, front, active_articulator, upper_teeth) %>%
  summary()
```


```{r}
# phonetic group distribution in association+image
features %>%
  pivot_longer(c("back", "active_articulator", "upper_teeth", "front", "soft_palate"),
               names_to = "feature",
               values_to = "value") %>%
  select(feature, value) %>%
  ggplot(
    aes(x=value,
        fill=feature,
        color=feature)
  ) +
  geom_histogram(position="dodge", bins=6)

```


```{r}
# check normality distribution using shapiro test
features %>%
  select(firstly_chosen_group, back, upper_teeth, active_articulator, front, soft_palate) %>%
  group_by(firstly_chosen_group) %>%
  apply(2, shapiro.test) # for majority of variables p-value < 0.05, ==> there are not normally distributed
```


# Tests


## main test

```
Since we are dealing with categorical variables and the size of samples is relativelty small, then it is a good idea to use Fisher Test:
H0: the last choice of word group is NOT influenced by choice of first word group
H1: the last choice of word group is influenced by choice of first word group
```


```{r}
fisher.test(data.tabs)
```
```
The p-value is > 0.05, which gives us a reason to accept h0. However, we should also check consonants and vowels independently
```


```{r}
# vowels
data.separated.vow_tab %>%
  fisher.test()
```


```{r}
# consonants
data.separated.cons_tab %>%
  fisher.test()
```


```
Vowels and consonants independently still have p-value > 0.0.5, which means that there are no statistically meaningfull reason to reject h0.
```



## support tests

```
Let's test support data
```



### Interest independence

```
H0: There is not relation between interest and word's category choices
H1: There is a relation between interest and word's category choices
```

```{r}
data %>%
  select(is_interested, firstly_chosen_group) %>%
  table() %>%
  fisher.test()
```


```{r}
data %>%
  select(is_interested, lastly_chosen_group) %>%
  table() %>%
  fisher.test()
```



```{r}
data.separated %>%
  select(is_interested, firstly_vow) %>%
  table() %>%
  fisher.test()
```


```{r}
data.separated %>%
  select(is_interested, lastly_vow) %>%
  table() %>%
  fisher.test()
```



```{r}
data.separated %>%
  select(is_interested, firstly_con) %>%
  table() %>%
  fisher.test()
```



```{r}
data.separated %>%
  select(is_interested, lastly_con) %>%
  table() %>%
  fisher.test()
```

```
For each test their p-value > 0.05, which means that there is not dependency between interest and phonetical combination
```


## association independence


```
Since more than a half of our phonetic groups are not normally distributed. Let's try to test their influency using generalised linear regression with Poisson ditribution
```


```{r}
glm_first_group <- 
  glm(firstly_chosen_group ~ active_articulator + upper_teeth + back + front + soft_palate, 
      data=features,
      family = poisson())
summary(glm_first_group)
```

```
It seems that intercept does all the work, which means that our phonems have no influenct on association and image.
```


```{r}
plot(cooks.distance(glm_first_group), type='b', pch = 16, col = "blue")
N <- nrow(features)
k <- 6
cutoff <- 1 / (N - k - 1)
abline(h=cutoff,lty=2)
```

```{r}
plot(glm_first_group, which=1)
```

# Conclusion

```
It seems that all connections between choices in our sample is a coincidance or there is not enough data to check statistically meaningfull connections between phonetical groups and commercial names
```
